---
title: "Darwin Core mapping"
subtitle: "Red List Butterflies Local List"
author:
- Lien Reyserhove
- Dimitri Brosens
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

This document describes how we map the checklist data to Darwin Core. 

# Setup

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Set locale (so we use UTF-8 character encoding):

```{r}
# This works on Mac OS X, might not work on other OS
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
```

Load libraries:

```{r}

# None core tidyverse packages:
library(magrittr)  # For %<>% pipes

# Other packages
library(tidyverse) # For data transformations
library(janitor)    # For cleaning input data
library(knitr)      # For nicer (kable) tables
library(readxl)     # To read excel files
library(stringr)    # To perform string operations
library(digest)     # To generate hashes
library(rgbif)      # To Match GBIF
library(assertable) # because it sneeded for rGBIF
library(inborutils) # wrap GBIF api data
library(dplyr)

```

Set file paths (all paths should be relative to this script):

```{r}
# Raw files:
raw_data_file = "../data/raw/dataset.txt"


# Processed files:
dwc_taxon_fileLoc = "../data/processed/dwc_checklistLoc/taxonLoc.csv"
dwc_distribution_file = "../data/processed/dwc_checklist/distribution.csv"

```

# Read and pre-process raw data

Create a data frame `raw_data` from the source data:

```{r}
# Read the source data:
raw_data <- read.csv("../data/raw/dataset.txt", sep = ";")
```

Clean the data somewhat: remove empty rows if present:

```{r}
raw_data %<>%
  remove_empty("rows") %>%    # Remove empty rows
  clean_names()               # Have sensible (lowercase) column names
```


## Further pre-processing:

Add prefix `raw_` to all column names to avoid name clashes with Darwin Core terms:

```{r}
colnames(raw_data) <- paste0("raw_", colnames(raw_data))
```
 
Preview data:

```{r}
head(raw_data, n = 100)
```

# Create taxonList core EU

```{r start_taxon}
taxonLoc <- raw_data
```

# taxonListEU

Create species list EU names

```{r}
taxonLoc %<>% select(raw_speciesname_local, raw_speciesname_europe) %>%
           distinct(raw_speciesname_local, .keep_all = TRUE) 
```

Preview data:

```{r}
head(taxonLoc , n = 100)
```

```{r}
# Vectorize the digest function (The digest() function isn't vectorized. So if you pass in a vector, you get one value for the whole vector rather than a digest for each element of the vector):
vdigest <- Vectorize(digest)

taxonLoc %<>% 

  gbif_species_name_match(name_col = "raw_speciesname_local", 
                          gbif_terms = c('usageKey', 'scientificName', 'rank', 
                                         'order', 'matchType','phylum', 
                                         'kingdom', 'genus','class',
                                         'confidence', 'synonym', 'status',
                                         'family'))%>%
  mutate(taxonID = paste(usageKey, vdigest (scientificName, algo="md5"), sep=":"))
```

Preview data:

```{r}
head(taxonLoc , n = 100)
View(taxonLoc)

```

## Add fixed terms mapping

Map the data to [Darwin Core Taxon](http://rs.gbif.org/core/dwc_taxon_2015-04-24.xml).
 
### language

```{r}
taxonLoc %<>% mutate(language = "en")
```


### license

```{r}
taxonLoc %<>% mutate(license = "http://creativecommons.org/publicdomain/zero/1.0/")

```

### rightsHolder

```{r}
taxonLoc %<>% mutate(rightsHolder = "INBO")  
```


### datasetID

```{r}
taxonLoc  %<>% mutate(datasetID = "to_be_completed")
```


### datasetName

```{r}
taxonLoc %<>% mutate(datasetName = "Red Lists Butterflies Europe_Locale")
```

### nomenclaturalCode

```{r}
taxonLoc %<>% mutate(nomenclaturalCode = "ICZN")
```



### Look at data

```{r}
head(taxonLoc)
View(taxonLoc)
```


 
## Post-processing

Remove the original columns:

```{r eval=FALSE, include=FALSE}
taxon %<>% select(-starts_with("raw_"))
```

Sort on `taxonID`:

```{r}
taxonLoc %<>% arrange(taxonID)
```

Preview data:

```{r}
head(taxonLoc)
```

Save to CSV:

```{r}
write.csv(taxonLoc, file = dwc_taxon_fileLoc, na = "", row.names = FALSE, fileEncoding = "UTF-8")
```

# Create distribution extension

```{r start_distribution}
distributionEU <- raw_data
```

## Term mapping

Map the data to [Species Distribution](http://rs.gbif.org/extension/gbif/1.0/distribution.xml).

### taxonID


### locationID
### locality
### countryCode
### occurrenceStatus
### establishmentMeans
### eventDate
### source

## Post-processing

Remove the original columns:

```{r}
distribution %<>% select(-starts_with("raw_"), -start_year, -end_year)
```

Sort on `taxonID`:

```{r}
distribution %<>% arrange(taxonID)
```

Preview data:

```{r}
distribution %>%
  mutate(source = substr(source, 1, 10)) %>% # Shorten source to make it easier to display
  head()
```

Save to CSV:

```{r}
write.csv(distribution, file = dwc_distribution_file, na = "", row.names = FALSE, fileEncoding = "UTF-8")
```

