---
title: "weighted_red_list_values_script"
author: "Dimitri Brosens"
date: "18 juli 2018"
output: html_document
---

This document describes how we map the checklist data to Darwin Core. 

# Setup

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Set locale (so we use UTF-8 character encoding):

```{r}
# This works on Mac OS X, might not work on other OS
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
```

Load libraries:

```{r}

library(tidyverse)        # For data transformations
library(magrittr)         # For %<>% pipes
library(googlesheets)     # To import and read Google spreadsheets 
library(janitor)          # For cleaning input data
library(digest)           # To generate hashes
library(knitr)
opts_chunk$set(echo = TRUE)
library(dplyr)
library(readxl)
library(RODBC)
library(gam)
```

Set file paths (all paths should be relative to this script):

```{r}
# Interim files:
taxa_regional_file = "../data/interim/taxa_regional.csv"
regional_threat_scores_file = "../data/interim/regional_threat_scores.csv"
country_info_file = "../data/interim/country_info.csv"

# Processed files:

weighted_red_list_values_file = "../data/processed/weighted_red_list_values.csv"

```

# Read raw data

The original spreadsheet can be found [here](https://docs.google.com/spreadsheets/d/1RvxpOYf2ZrTu9nsTLumoi-G-GGhh6_lV37TNtPiVES4/edit?ts=5b1f937f#gid=1323544344). We need to retrieve this spreadsheet and select the specific Googlesheet:

```{r}
retrieve_spreadsheet <- gs_title("EuroButt")
```

This spreadsheet contains four relevant different worksheets: `taxa_regional`, `regional_threat_scores`, `country_info` and `references`. We select these here and save them as separate data frames. We will then continue with the `taxa_regional` as this is the original dataframe.

```{r}
taxa_regional <- retrieve_spreadsheet %>% gs_read("taxa_regional") 
regional_threat_scores <- retrieve_spreadsheet %>% gs_read("regional_threat_scores")
country_info <- retrieve_spreadsheet %>% gs_read("country_info")

```

Create a data frame `raw_data` from `taxa_regional`, which is the original dataframe and the basis for the calculation of the values:

```{r}
raw_data <- taxa_regional
```

### Pre-processing

Clean data somewhat

```{r}
raw_data %<>%
  remove_empty("rows") %>%    # Remove empty rows
  clean_names()               # Have sensible (lowercase) column names
```

Add prefix `raw_` to all column names to avoid name clashes with Darwin Core terms:

```{r}
colnames(raw_data) <- paste0("raw_", colnames(raw_data))
```

Select columns for which `raw
-use = Y`:

```{r}
raw_data %<>% filter(raw_use == 'y') 
```

Preview data:

```{r}
head(raw_data, n = 100)
```



```{r}
raw_data %<>%
  select(raw_speciesname_europe, raw_country_code, raw_red_list_category)%>%    
  group_by(raw_speciesname_europe, raw_country_code,raw_red_list_category, add = FALSE) %>%
  summarise()
head(raw_data, n=50)
View(raw_data)
```
 

```{r}
raw_data %<>%
  group_by(raw_speciesname_europe)
View(raw_data)
```


